Detecting the impact of land cover change on observed rainfall.

### 2019-03-14 Manuscripts/PeerJRainfall_Landuse/Rainfall_Landuse.Rmd
Update files for change to png
Upload since push does not work?

---
title: Detecting the impact of land cover change on observed rainfall.
preprint: false
author:
  - name: Chun X. Liang
    affiliation: 1
  - name: Floris F. van Ogtrop
    affiliation: 1
  - name: R. Willem Vervoort
    affiliation: 1
    corresponding: true
    email: willem.vervoort@sydney.edu.au
affiliation:
  - code: 1
    address: Sydney Institute of Agriculture, The University of Sydney, NSW 2006
abstract: >
 Analysis of observational data to pinpoint impact of land cover change on local rainfall are difficult due to multiple environmental factors that cannot be strictly controlled. In this study we use a statistical approach to identify the relationship between removal of tree cover and rainfall with data from best available sources for two large areas in Australia. Gridded rainfall data between 1979 and 2015 was used for the areas, while large scale (exogenous) effects were represented by mean rainfall across a much larger area and climatic indicators, such as SOI and IOD. Both general additive modelling and step trend tests were used for the analysis. 
  For a region in south central Queensland, the reported change in tree clearing between 2002 - 2005 did not result in strong statistically significant precipitation changes. On the other hand, results from a bushfire affected region on the border of New SOuth Wales and Victoria suggests significant changes in the rainfall due to changes in tree cover. This indicates the method works better when a abrupt change in the data can be clearly identified. The results from the step trend test also mainly identified a positive relationship between the tree cover and the rainfall at p < 0.1 at the NSW/Victoria region. High rainfall variability and possible regrowth could have impacted the results in the Queensland region.
header-includes: >
  \usepackage{lipsum}
  \usepackage{textcomp, rotating}
bibliography: reflist_chap3.bib
date: "`r Sys.Date()`"
output: 
  rticles::peerj_article:
    base_format: bookdown::pdf_document2 # bookdown::pdf_document2 # for using \@ref()
---
```{r packages, echo=F, warnings=F, message=F, eval=T}
suppressWarnings(require(knitr))
suppressWarnings(require(bookdown))
suppressWarnings(require(pander))
suppressWarnings(library(fields))
suppressWarnings(library(graphics))
suppressWarnings(library(maptools))
suppressWarnings(library(plotKML))
suppressWarnings(library(raster))
suppressWarnings(library(reshape))
suppressWarnings(library(rgeos))
suppressWarnings(library(rgdal))
suppressWarnings(library(RColorBrewer))
suppressWarnings(library(tidyverse))
suppressWarnings(require(lubridate))
suppressWarnings(library(deseasonalize))
suppressWarnings(library(mgcv))
suppressWarnings(library(ncdf4))
suppressWarnings(require(gridExtra))
suppressWarnings(require(kableExtra))
#library(nnet)
```

@@

include_graphics("figures/map_selreg.jpg")
```

@@

```{r bushfire, fig.cap = "Location of bushfires occurring in January 2003, in and around the NSW/VIC study region, as shown by the red pixels. The map shows the large area in the Kosciuszko national park that has been burned. Some locations in the southwest of the Australian Capital Territory (ACT) have also experienced intensive bushfires.", out.width="90%", echo=F}
include_graphics("figures/bushfire_nswvic.jpg")
```

@@

```{r figure3-tc-simple, fig.cap= "The expected evolution of the land surface after trees have been removed in (a) the QLD region and (b) the NSW/VIC region.", out.width="90%", echo=F}
include_graphics("figures/tc_simple.jpg")
```

@@

jpeg("../../figures/bushfire_nswvic.jpg", width=960, height=960)
  par(mar=c(3,3,1,6),xpd=T)
  plot(e222,border="white")  
  image(fire_loc,add=T, breaks=c(0,0.99,366.5,10000),col=c("white","red","white"))
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
#   text(1430000,-1000000,"study region",cex=1.2)
  text(149.1, -35.3,"ACT",font=2,cex=1.2)
  text(148.4, -36,"Kosciuszko",font=2,cex=0.9)#,srt=90)
  text(147.5, -36.4,"VIC",font=2,cex=1.2)
  text(148, -35.5,"NSW",font=2,cex=1.2)
  image.plot(zlim=c(0,1),breaks=c(0,0.5,1),lab.breaks=c("","burned \narea",""),legend.only=T,horizontal=F,col=c("red","red"),legend.shrink=0.03,legend.width=1.3,legend.args=list(text="",line=0),lwd.poly=0.1)
legend(149,-36.1,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=0.9)
 dev.off()

@@

jpeg("../../figures/slats.jpg", width=1024, height = 760)
Data %>% 
  gather(key="Year",value="clearing", `1988-1989`:`2014-2015`) %>%
  filter(Bioregion == "Mulga Lands" |
           Bioregion == "Brigalow Belt") %>%
  ggplot(aes(Year,clearing, fill=Bioregion)) + 
  geom_bar(stat="identity") + ylab("Clearing (ha)") +
  scale_fill_grey() +
  theme(axis.text.x = element_text(size = rel(2),
                                   angle = 45, vjust=0.5),
        axis.text.y = element_text(size = rel(2)),
        axis.title = element_text(size = rel(2), face = "bold"),
        legend.text = element_text(size = rel(2)),
        legend.title = element_text(size = rel(2), face = "bold"))
dev.off()

@@

jpeg("../../figures/tc_figs.jpg", width=1280, height=640, quality=100)
par(mfrow=c(1,2), mar=c(5,3,5,5))

@@

include_graphics("figures/Rainfall_resid.jpg")
```

@@

jpeg("../../figures/hist_rainfall.jpg")
rain %>%
  ggplot(aes(Rainfall)) + geom_histogram(fill="blue") +
  facet_wrap(~State) + xlab("Monthly Rainfall (mm)") + 
  ylab("Frequency") +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

jpeg("../../figures/autocorrelation.jpg")
acf_df %>%
  gather(key="acf", value="cor", acf_Qld:pacf_NSW) %>%
  ggplot(aes(lag, cor)) + geom_bar(stat="identity", fill = "blue") +
  facet_wrap(~acf) + xlab("Lag") + ylab("correlation")
dev.off()

@@

jpeg("../../figures/Rainfall_resid.jpg", width=960, height=960)
plot_df %>%
  ggplot(aes(Dates,Resid)) + geom_line() + 
  geom_vline(xintercept = as.Date("2003-01-01"), 
             col="red", size=1.5) +
  geom_text(aes(x=as.Date("2004-12-01"),y=150),
            label="2003", col="red", size=7) +
  facet_wrap(~State, ncol = 1) + xlab("Date") +
  ylab("Deseasonalised rainfall for the NSW and Qld locations") +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

jpeg("../../figures/cor_qld.jpg", width=960, height=960)
plot_df_qld %>%
  gather(key="SST", value="correlation", SOI:IOD) %>%
  ggplot(aes(Lag,correlation)) + 
  geom_bar(fill="blue", stat="identity") +
  facet_wrap(~SST) + xlab("Lag in months") + ylab("correlation") +
  xlim(c(-12.5,0.5)) +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

jpeg("../../figures/cor_nswvic.jpg", width=960, height=960)
plot_df %>%
  gather(key="SST", value="correlation", SOI:IOD) %>%
  ggplot(aes(Lag,correlation)) + 
  geom_bar(fill="blue", stat="identity") +
  facet_wrap(~SST) + xlab("Lag in months") + ylab("correlation") +
  xlim(c(-12.5,0.5)) +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

```{r cor-rain-qld, fig.cap="Cross-correlation of six climate indices and rainfall in QLD study region. For the PDO analysis, 108-year rainfall data (1900 - 2008) are used. Otherwise, 36-year rainfall data are used. The correlation with NINO 3 is not shown as it is very similar to but weaker than for NINO 3.4.", out.width="90%", echo=F}
include_graphics("figures/cor_qld.tif")
```

```{r cor-rain-nsw, fig.cap="Cross-correlation of six climate indices and rainfall in NSW/VIC study region.", out.width="90%", echo=F}
include_graphics("figures/cor_nswvic.tif")
```

@@

```{r hist-rain, fig.cap="Distribution of monthly rainfall in (a) QLD and (b) NSW/VIC. Using a Kolmogorov-Smirnov test with shape = 1 and 2.4  for QLD and NSW/VIC respectively, rainfall in both regions can be modelled as a gamma distribution.", echo=F, out.width="90%"}
include_graphics("figures/hist_rainfall.jpg")
```

@@

jpeg("../figures/tc_simple.jpg", width=720, height=480)
par(mar=c(5,6,1,1), mfrow=c(1,2))
plot(data1$year,data1$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",xaxt="n",yaxt="n")
axis(1,at=c(1,2,3),labels=c("before","LCC","after"),cex.axis=1.3)
axis(2,at=c(1,2),label=c("removed","tree"),cex.axis=1.3,las=2)
par(mar=c(5,6,1,1))
plot(data2$year,data2$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",xaxt="n",yaxt="n")
axis(1,at=c(1,2,4),labels=c("before","LCC","after"),cex.axis=1.3)
axis(2,at=c(1,2),label=c("removed","tree"),cex.axis=1.3,las=2)
dev.off()

@@

jpeg("../../figures/hist_sign_LC.jpg")
par(mfrow=c(2,1))
hist(unlist(lapply(1:length(coef.rain30.nswvic),function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",summary(coef.rain30.nswvic[[x]])$p.coeff[3],NA))), xlab = "value of tree cover estimate", main="NSW full data period")
hist(unlist(lapply(1:length(coef.rain30.qld4),function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",summary(coef.rain30.qld4[[x]])$p.coeff[3],NA))), xlab = "value of tree cover estimate", main="Qld full data period")
dev.off()

@@

# Now make the final plot for the paper with both sites and 2009 and all data
  jpeg("../../figures/Cp_30yrs.jpg", width=1440, height=720)
  par(mfrow=c(1,2))
  par(mar=c(5,5,3,7))
  # QLD first
  plot(e111,xlab="",ylab="", cex.lab = 2,
       cex.axis=2)
  image(dist_Cp30_qld4,col=c("red","orange","yellow","light blue"),breaks=c(0,0.05,0.1,0.2,1),legend=F,add=T)
#   plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=5,add=T)
  plot(e111,add=T)
legend(148,-27.5,c("State","MDB"),col=c("grey","grey"),lwd=c(2,5),bty="n",title="Boundary",cex=1.5)

@@

# plotting
p_plot_df <- rbind(plinear_qld,plinear_nsw)
p_plot_df %>%
  gather(key="Variable", value="p_value", MDB:trd) %>%
  ggplot(aes(p_value,fill=state)) + geom_histogram(alpha=0.5) +
  facet_wrap(~Variable) +
  scale_fill_manual(values = c("blue","red")) +
  theme(axis.title = element_text(size = rel(2)),
        axis.text = element_text(size=rel(1)),
        strip.text = element_text(face="bold", size= rel(2)),
        legend.text = element_text(face="bold", size= rel(1)),
        legend.title = element_text(face="bold", size= rel(2)))
# produce also as jpeg
jpeg("../../figures/pvalues_linear_fullmodel.jpg", height=960, width=960)
p_plot_df %>%
  gather(key="Variable", value="p_value", MDB:trd) %>%
  ggplot(aes(p_value,fill=state)) + geom_histogram(alpha=0.5) +
  facet_wrap(~Variable) +
  scale_fill_manual(values = c("blue","red")) +
  theme(axis.title = element_text(size = rel(2)),
        axis.text = element_text(size=rel(1)),
        strip.text = element_text(face="bold", size= rel(2)),
        legend.text = element_text(face="bold", size= rel(1)),
        legend.title = element_text(face="bold", size= rel(2)))
dev.off()

@@

p_plot_df %>%
  gather(key="Variable", value="p_value", SOI:NINO4) %>%
  ggplot(aes(p_value,fill=state)) + geom_histogram(alpha=0.5) +
  facet_wrap(~Variable) +
  scale_fill_manual(values = c("blue","red")) +
  theme(axis.title = element_text(size = rel(2)),
        axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(face="bold", size= rel(2)),
        legend.text = element_text(face="bold", size= rel(1.5)),
        legend.title = element_text(face="bold", size= rel(2)))
# produce also as jpg
jpeg("../../figures/pvalues_fullmodel.jpg", height=960, width=960)
p_plot_df %>%
  gather(key="Variable", value="p_value", SOI:NINO4) %>%
  ggplot(aes(p_value,fill=state)) + geom_histogram(alpha=0.5) +
  facet_wrap(~Variable) +
  scale_fill_manual(values = c("blue","red")) +
  theme(axis.title = element_text(size = rel(2)),
        axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(face="bold", size= rel(2)),
        legend.text = element_text(face="bold", size= rel(1.5)),
        legend.title = element_text(face="bold", size= rel(2)))
dev.off()
rm(p_plot_df)

@@

jpeg("../../figures/rsq_fullmodel.jpg", height=480, width=480)
plot_df %>%
  ggplot(aes(`r.squared`*100, fill=state)) + geom_histogram() +
  scale_fill_manual(values = c("blue","red")) +
  xlab(expression(paste("% of variability explained (",r^2,")")))
dev.off()

@@

## plots
jpeg("../../figures/GAMresidualTime.jpeg", height=960, width=960)
residuals %>%
  gather(key="State", value="residual", QLD, NSWVIC) %>%
  ggplot(aes(Dates,residual)) + geom_line(colour="blue") +
  geom_vline(xintercept=as.numeric(as.Date("2003-01-01")), colour="red",
             size=1.5) +
  facet_wrap(~State)  +
  theme(axis.text = element_text(size=rel(1.5)),
        axis.title = element_text(size=rel(2)),
        strip.text = element_text(size=rel(2), face = "bold"))
dev.off()

#
#
## plot gam.check
jpeg("../../figures/gam_check.jpg", width=720, height=960)
  par(mfrow=c(4,2))
  gam.check(coef.rain30.qld4[[1]])
  gam.check(coef.rain30.nswvic[[1]])
dev.off()

@@

plot_df <- rbind(qld_df, nsw_df)
jpeg("../../figures/pmodel_explHistogram.jpg")
plot_df %>%
  ggplot(aes(rsq)) + geom_histogram(fill="blue") +
  facet_wrap(~state) +
  xlab(expression(paste("% of variability explained (",r^2,")"))) +
  theme(axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(size=rel(2)),
        axis.title = element_text(size=rel(2)))
 dev.off()

@@

jpeg("../../figures/ResidualBoxplot.jpg")
resplot_df %>%
  filter(Date >= ymd("1998-01-01")) %>%
  ggplot(aes(Date,residuals, fill=State)) +
  geom_boxplot(aes(group=year(Date))) +
  facet_wrap(~State, ncol=1)
dev.off()

# residuals before and after change
#par(mfrow=c(1,2))
jpeg("figures/ResidualBoxplotchange.jpg")
resplot_df %>%
  filter(Date >= ymd("1998-01-01")) %>%
  mutate(Change = ifelse(year(Date) > 2003, "After", "Before")) %>%
  ggplot(aes(Change,residuals, fill= State)) +
  geom_boxplot(aes(group=Change)) + facet_wrap(~State, ncol=1) +
  guides(fill=FALSE) + ggtitle("whole period")
dev.off()

@@

plot_df <- rbind(qld_df, nsw_df)
jpeg("../../figures/soi_explHistogram.jpg", width=960, height=960)
plot_df %>%
  ggplot(aes(rsq)) + geom_histogram(fill="blue") +
  facet_wrap(~state) +
  xlab(expression(paste("% of variability explained (adj",r^2,")"))) +
  theme(axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(size=rel(2)),
        axis.title = element_text(size=rel(2)))
 dev.off()

@@

```{r Zscore, echo=F}
table2 <- data_frame(" "= c("$Z'>0$  and rainfall decreases after change",
    "$Z'<0$ and rainfall increases post change",
    "$Z'=0$ and rainfall does not change"))
kable(table2, caption="The interpretation of Z' score in the step trend test",
      booktabs=T)
```

@@

```{r rsqmodel, fig.cap = "The distribution of the variance explained (adjusted $r^2$ ) by the full regression model", echo=F, out.width="90%"}
include_graphics("figures/rsq_fullmodel.jpg")

@@

```{r residuals, fig.cap ="The residual analysis of a sample pixel in the QLD region (top) and NSW/VIC region (bottom).", echo=F, out.width="90%"}
include_graphics("figures/gam_check.jpg")
```

>>>>

---
title: Detecting the impact of land cover change on observed rainfall.
preprint: false
author:
  - name: Chun X. Liang
    affiliation: 1
  - name: Floris F. van Ogtrop
    affiliation: 1
  - name: R. Willem Vervoort
    affiliation: 1
    corresponding: true
    email: willem.vervoort@sydney.edu.au
affiliation:
  - code: 1
    address: Sydney Institute of Agriculture, The University of Sydney, NSW 2006
abstract: >
 Analysis of observational data to pinpoint impact of land cover change on local rainfall are difficult due to multiple environmental factors that cannot be strictly controlled. In this study we use a statistical approach to identify the relationship between removal of tree cover and rainfall with data from best available sources for two large areas in Australia. Gridded rainfall data between 1979 and 2015 was used for the areas, while large scale (exogenous) effects were represented by mean rainfall across a much larger area and climatic indicators, such as Southern Oscillation Index and Indian Ocean Dipole. Both general additive modelling and step trend tests were used for the analysis. 
  For a region in south central Queensland, the reported change in tree clearing between 2002 - 2005 did not result in strong statistically significant precipitation changes. On the other hand, results from a bushfire affected region on the border of New South Wales and Victoria suggests significant changes in the rainfall due to changes in tree cover. This indicates the method works better when a abrupt change in the data can be clearly identified. The results from the step trend test also mainly identified a positive relationship between the tree cover and the rainfall at p < 0.1 at the NSW/Victoria region. High rainfall variability and possible regrowth could have impacted the results in the Queensland region.
header-includes: >
  \usepackage{lipsum}
  \usepackage{textcomp, rotating}
bibliography: reflist_chap3.bib
date: "`r Sys.Date()`"
output: 
  rticles::peerj_article:
    base_format: bookdown::pdf_document2 # bookdown::pdf_document2 # for using \@ref()
---
```{r packages, echo=F, warnings=F, message=F, eval=T}
suppressWarnings(require(knitr))
suppressWarnings(require(bookdown))
suppressWarnings(require(pander))
suppressWarnings(library(fields))
suppressWarnings(library(graphics))
suppressWarnings(library(maptools))
suppressWarnings(library(plotKML))
suppressWarnings(library(raster))
suppressWarnings(library(reshape))
suppressWarnings(library(rgeos))
suppressWarnings(library(rgdal))
suppressWarnings(library(RColorBrewer))
suppressWarnings(library(tidyverse))
suppressWarnings(require(lubridate))
suppressWarnings(library(deseasonalize))
suppressWarnings(library(mgcv))
suppressWarnings(library(ncdf4))
suppressWarnings(require(gridExtra))
suppressWarnings(require(kableExtra))
#library(nnet)
```

@@

include_graphics("figures/map_selreg.png")
```

@@

```{r bushfire, fig.cap = "Location of bushfires occurring in January 2003, in and around the NSW/VIC study region, as shown by the red pixels. The map shows the large area in the Kosciuszko national park that has been burned. Some locations in the southwest of the Australian Capital Territory (ACT) have also experienced intensive bushfires.", out.width="90%", echo=F}
include_graphics("figures/bushfire_nswvic.png")
```

@@

```{r figure3-tc-simple, fig.cap= "The expected evolution of the land surface after trees have been removed in (a) the QLD region and (b) the NSW/VIC region.", out.width="90%", echo=F}
include_graphics("figures/tc_simple.png")
```

@@

png("figures/bushfire_nswvic.png", width=960, height=960)
  par(mar=c(3,3,1,7),xpd=T)
  plot(e222,border="white", xlab="",ylab="", cex.axis=2)  
  image(fire_loc,add=T, breaks=c(0,0.99,366.5,10000),col=c("white","red","white"))
  plot(e222,add=T)
  plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=3,add=T)
#   text(1430000,-1000000,"study region",cex=1.2)
  text(149.1, -35.3,"ACT",font=2,cex=2)
  text(148.4, -36,"Kosciuszko",font=2,cex=2)#,srt=90)
  text(147.5, -36.4,"VIC",font=2,cex=2)
  text(148, -35.5,"NSW",font=2,cex=2)
  image.plot(zlim=c(0,1),breaks=c(0,0.5,1),lab.breaks=c("","burned \narea",""),axis.args=list(cex.axis=1.5),legend.only=T,horizontal=F,col=c("red","red"),legend.shrink=0.07,legend.width=1.0,lwd.poly=0.1, legend.mar=7.1)
legend(148.7,-36.2,c("State","MDB"),col=c("grey","grey"),lwd=c(1,3),bty="n",title="Boundary",cex=1.5)
 dev.off()

@@

png("figures/slats.png", width=1200, height = 900)
Data %>% 
  gather(key="Year",value="clearing", `1988-1989`:`2014-2015`) %>%
  filter(Bioregion == "Mulga Lands" |
           Bioregion == "Brigalow Belt") %>%
  ggplot(aes(Year,clearing, fill=Bioregion)) + 
  geom_bar(stat="identity") + ylab("Clearing (ha)") +
  scale_fill_grey() +
  theme(axis.text.x = element_text(size = rel(2),
                                   angle = 45, vjust=0.5),
        axis.text.y = element_text(size = rel(2)),
        axis.title = element_text(size = rel(2), face = "bold"),
        legend.text = element_text(size = rel(2)),
        legend.title = element_text(size = rel(2), face = "bold"))
dev.off()

@@

png("figures/tc_figs.png", width=1500, height=900)
par(mfrow=c(1,2), mar=c(5,3,5,5))

@@

include_graphics("figures/Rainfall_resid.png")
```

@@

png("figures/hist_rainfall.png", width=900, height=900)
rain %>%
  ggplot(aes(Rainfall)) + geom_histogram(fill="blue") +
  facet_wrap(~State) + xlab("Monthly Rainfall (mm)") + 
  ylab("Frequency") +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

png("figures/autocorrelation.png", width =900, height=900)
acf_df %>%
  gather(key="acf", value="cor", acf_Qld:pacf_NSW) %>%
  ggplot(aes(lag, cor)) + geom_bar(stat="identity", fill = "blue") +
  facet_wrap(~acf) + xlab("Lag") + ylab("correlation")
dev.off()

@@

png("figures/Rainfall_resid.png", width=960, height=960)
plot_df %>%
  ggplot(aes(Dates,Resid)) + geom_line() + 
  geom_vline(xintercept = as.Date("2003-01-01"), 
             col="red", size=1.5) +
  geom_text(aes(x=as.Date("2004-12-01"),y=150),
            label="2003", col="red", size=7) +
  facet_wrap(~State, ncol = 1) + xlab("Date") +
  ylab("Deseasonalised rainfall for the NSW and Qld locations") +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

png("figures/cor_qld.png", width=960, height=960)
plot_df_qld %>%
  gather(key="SST", value="correlation", SOI:IOD) %>%
  ggplot(aes(Lag,correlation)) + 
  geom_bar(fill="blue", stat="identity") +
  facet_wrap(~SST) + xlab("Lag in months") + ylab("correlation") +
  xlim(c(-12.5,0.5)) +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

png("figures/cor_nswvic.png", width=960, height=960)
plot_df %>%
  gather(key="SST", value="correlation", SOI:IOD) %>%
  ggplot(aes(Lag,correlation)) + 
  geom_bar(fill="blue", stat="identity") +
  facet_wrap(~SST) + xlab("Lag in months") + ylab("correlation") +
  xlim(c(-12.5,0.5)) +
  theme(axis.text=element_text(size=rel(2)),
        axis.title=element_text(size=rel(2.5)),
        strip.text =element_text(size=rel(2), face="bold"))
dev.off()

@@

```{r cor-rain-qld, fig.cap="Cross-correlation of six climate indices and rainfall in QLD study region. For the PDO analysis, 108-year rainfall data (1900 - 2008) are used. Otherwise, 36-year rainfall data are used. The correlation with NINO 3 is not shown as it is very similar to but weaker than for NINO 3.4.", out.width="90%", echo=F}
include_graphics("figures/cor_qld.png")
```

```{r cor-rain-nsw, fig.cap="Cross-correlation of six climate indices and rainfall in NSW/VIC study region.", out.width="90%", echo=F}
include_graphics("figures/cor_nswvic.png")
```

@@

```{r hist-rain, fig.cap="Distribution of monthly rainfall in (a) QLD and (b) NSW/VIC. Using a Kolmogorov-Smirnov test with shape = 1 and 2.4  for QLD and NSW/VIC respectively, rainfall in both regions can be modelled as a gamma distribution.", echo=F, out.width="90%"}
include_graphics("figures/hist_rainfall.png")
```

@@

png("figures/tc_simple.png", width=1200, height=900)
par(mar=c(5,8,4,3), mfrow=c(1,2))
plot(data1$year,data1$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",xaxt="n",yaxt="n")
axis(1,at=c(1,2,3),labels=c("before","LCC","after"),cex.axis=2, cex.lab=2)
axis(2,at=c(1,2),label=c("removed","tree"),cex.axis=2,las=2, cex.lab=1.5)
par(mar=c(5,8,4,3))
plot(data2$year,data2$tc,type="l",ylim=c(0.5,2.5),xlab="",ylab="",xaxt="n",yaxt="n")
axis(1,at=c(1,2,4),labels=c("before","LCC","after"),cex.axis=2, cex.lab=2)
axis(2,at=c(1,2),label=c("removed","tree"),cex.axis=2, cex.lab=2,las=1.5)
dev.off()

@@

png("figures/hist_sign_LC.png", width=900, height = 900)
par(mfrow=c(2,1))
hist(unlist(lapply(1:length(coef.rain30.nswvic),function(x) ifelse(summary(coef.rain30.nswvic)[x,2]=="gam",summary(coef.rain30.nswvic[[x]])$p.coeff[3],NA))), xlab = "value of tree cover estimate", main="NSW full data period")
hist(unlist(lapply(1:length(coef.rain30.qld4),function(x) ifelse(summary(coef.rain30.qld4)[x,2]=="gam",summary(coef.rain30.qld4[[x]])$p.coeff[3],NA))), xlab = "value of tree cover estimate", main="Qld full data period")
dev.off()

@@

# Now make the final plot for the paper with both sites and 2009 and all data
  png("figures/Cp_30yrs.png", width=1800, height=900)
  par(mfrow=c(1,2))
  par(mar=c(5,5,3,7))
  # QLD first
  plot(e111,xlab="",ylab="", cex.lab = 2,
       cex.axis=2)
  image(dist_Cp30_qld4,col=c("red","orange","yellow","light blue"),breaks=c(0,0.05,0.1,0.2,1),legend=F,add=T)
#   plot(aus_reg2,border="grey",add=T)
  plot(mdb_reg2,border="grey",lwd=5,add=T)
  plot(e111,add=T)
legend(148,-27.5,c("State","MDB"),col=c("grey","grey"),lwd=c(2,5),bty="n",title="Boundary",cex=1.5)

@@

# plotting
p_plot_df <- rbind(plinear_qld,plinear_nsw)
p_plot_df %>%
  gather(key="Variable", value="p_value", MDB:trd) %>%
  ggplot(aes(p_value,fill=state)) + geom_histogram(alpha=0.5) +
  facet_wrap(~Variable) +
  scale_fill_manual(values = c("blue","red")) +
  theme(axis.title = element_text(size = rel(2)),
        axis.text = element_text(size=rel(1)),
        strip.text = element_text(face="bold", size= rel(2)),
        legend.text = element_text(face="bold", size= rel(1)),
        legend.title = element_text(face="bold", size= rel(2)))
# produce also as jpeg
png("figures/pvalues_linear_fullmodel.png", height=960, width=960)
p_plot_df %>%
  gather(key="Variable", value="p_value", MDB:trd) %>%
  ggplot(aes(p_value,fill=state)) + geom_histogram(alpha=0.5) +
  facet_wrap(~Variable) +
  scale_fill_manual(values = c("blue","red")) +
  theme(axis.title = element_text(size = rel(2)),
        axis.text = element_text(size=rel(1)),
        strip.text = element_text(face="bold", size= rel(2)),
        legend.text = element_text(face="bold", size= rel(1)),
        legend.title = element_text(face="bold", size= rel(2)))
dev.off()

@@

p_plot_df %>%
  gather(key="Variable", value="p_value", SOI:NINO4) %>%
  ggplot(aes(p_value,fill=state)) + geom_histogram(alpha=0.5) +
  facet_wrap(~Variable) +
  scale_fill_manual(values = c("blue","red")) +
  theme(axis.title = element_text(size = rel(2)),
        axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(face="bold", size= rel(2)),
        legend.text = element_text(face="bold", size= rel(1.5)),
        legend.title = element_text(face="bold", size= rel(2)))
# produce also as jpg
png("figures/pvalues_fullmodel.png", height=960, width=960)
p_plot_df %>%
  gather(key="Variable", value="p_value", SOI:NINO4) %>%
  ggplot(aes(p_value,fill=state)) + geom_histogram(alpha=0.5) +
  facet_wrap(~Variable) +
  scale_fill_manual(values = c("blue","red")) +
  theme(axis.title = element_text(size = rel(2)),
        axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(face="bold", size= rel(2)),
        legend.text = element_text(face="bold", size= rel(1.5)),
        legend.title = element_text(face="bold", size= rel(2)))
dev.off()
rm(p_plot_df)

@@

png("figures/rsq_fullmodel.png", height=900, width=900)
plot_df %>%
  ggplot(aes(`r.squared`*100, fill=state)) + geom_histogram() +
  scale_fill_manual(values = c("blue","red")) +
  xlab(expression(paste("% of variability explained (",r^2,")"))) +
    theme(axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(size=rel(2)),
        axis.title = element_text(size=rel(2)))
dev.off()

@@

## plots
png("figures/GAMresidualTime.png", height=960, width=960)
residuals %>%
  gather(key="State", value="residual", QLD, NSWVIC) %>%
  ggplot(aes(Dates,residual)) + geom_line(colour="blue") +
  geom_vline(xintercept=as.numeric(as.Date("2003-01-01")), colour="red",
             size=1.5) +
  facet_wrap(~State)  +
  theme(axis.text = element_text(size=rel(1.5)),
        axis.title = element_text(size=rel(2)),
        strip.text = element_text(size=rel(2), face = "bold"))
dev.off()

#
#
## plot gam.check
png("figures/gam_check.png", width=960, height=960)
  par(mfrow=c(4,2))
  gam.check(coef.rain30.qld4[[1]], cex=1.5, lwd=2,cex.lab=2, cex.axis = 1.5)
  gam.check(coef.rain30.nswvic[[1]], cex=1.5, lwd=2,cex.lab=2, cex.axis = 1.5)
dev.off()

@@

plot_df <- rbind(qld_df, nsw_df)
png("figures/pmodel_explHistogram.png", width=960, height=960)
plot_df %>%
  ggplot(aes(rsq)) + geom_histogram(fill="blue") +
  facet_wrap(~state) +
  xlab(expression(paste("% of variability explained (",r^2,")"))) +
  theme(axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(size=rel(2)),
        axis.title = element_text(size=rel(2)))
 dev.off()

@@

png("figures/ResidualBoxplot.png", width=960, height=960)
resplot_df %>%
  filter(Date >= ymd("1998-01-01")) %>%
  ggplot(aes(Date,residuals, fill=State)) +
  geom_boxplot(aes(group=year(Date))) +
  facet_wrap(~State, ncol=1) +
    theme(axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(size=rel(2)),
        axis.title = element_text(size=rel(2)))
dev.off()

# residuals before and after change
#par(mfrow=c(1,2))
png("figures/ResidualBoxplotchange.png", width=960, height=960)
resplot_df %>%
  filter(Date >= ymd("1998-01-01")) %>%
  mutate(Change = ifelse(year(Date) > 2003, "After", "Before")) %>%
  ggplot(aes(Change,residuals, fill= State)) +
  geom_boxplot(aes(group=Change)) + facet_wrap(~State, ncol=1) +
  guides(fill=FALSE) +
  theme(axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(size=rel(2)),
        axis.title = element_text(size=rel(2)))
dev.off()

@@

plot_df <- rbind(qld_df, nsw_df)
png("figures/soi_explHistogram.png", width=960, height=960)
plot_df %>%
  ggplot(aes(rsq)) + geom_histogram(fill="blue") +
  facet_wrap(~state) +
  xlab(expression(paste("% of variability explained (adj",r^2,")"))) +
  theme(axis.text = element_text(size=rel(1.5)),
        strip.text = element_text(size=rel(2)),
        axis.title = element_text(size=rel(2)))
 dev.off()

@@

```{r Zscore, echo=F}
table2 <- data_frame(" "= c("$Z'>0$  and rainfall decreases after change",
    "$Z'<0$ and rainfall increases post change",
    "$Z'=0$ and rainfall does not change"))
kable(table2, 'latex',
      caption="The interpretation of Z' score in the step trend test",
      booktabs=T, escape=F)
```

@@

```{r rsqmodel, fig.cap = "The distribution of the variance explained (adjusted $r^2$ ) by the full regression model", echo=F, out.width="90%"}
include_graphics("figures/rsq_fullmodel.png")

@@

```{r residuals, fig.cap ="The residual analysis of a sample pixel in the QLD region (top) and NSW/VIC region (bottom).", echo=F, out.width="90%"}
include_graphics("figures/gam_check.png")
```

```{r rsq, fig.cap = "The performance of the regression model if rainfall is only modelled by the climate drivers. It shows the percentage of rainfall variability that can be explained by the climate drivers for the Qld and NSW/VIC region", echo=F, out.width="70%"}
include_graphics("figures/soi_explhistogram.png")
